---
title: "Hands-on_Ex03"
author: "Chen Jinghui"
date: "Sep 3 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  message: false #no warning messages go out
  freeze: true  #will not re-render all items
---

[**7 Network Constrained Spatial Point Patterns Analysis**]{.underline}

7.3 Installing and launching the R packages

```{r}
pacman::p_load(sf, spNetwork, tmap, tidyverse)
```

7.4 Data Import and Preparation

```{r}
network <- st_read(dsn="C:/jhui-chen/ISSS626-GAA/Hands-on_Ex/Hands-on_Ex03/data/geospatial",
                   layer="Punggol_St")%>%
st_transform(crs = 3414)
st_crs(st_geometry(network)) 

#plot(st_geometry(network))
```

```{r}
childcare <- st_read(dsn="C:/jhui-chen/ISSS626-GAA/Hands-on_Ex/Hands-on_Ex03/data/geospatial",
                     layer="Punggol_CC")%>%
st_transform(crs = 3414)%>%
  st_zm()  # remove Z (elevation) and M (measure) dimensions from spatial data. 
st_crs(childcare)
#plot(st_geometry(childcare))
```

```{r}
childcare
```

```{r}
network
```

7.5 Visualising the Geospatial Data

```{r}
plot(st_geometry(network))
plot(childcare,add=T,col='red',pch = 19)
```

```{r}
tmap_mode('view')
tm_shape(childcare) +
  tm_dots() +
  tm_shape(network) +
  tm_lines()

tmap_mode("plot")
```

7.6 Network KDE (NKDE) Analysis

7.6.1 Preparing the lixels objects

```{r}
lixels <- lixelize_lines(network,
                         700,
                         mindist = 375)
#plot(lixels)
#view(lixels)
```

7.6.2 Generating line centre points

```{r}
samples <- lines_center(lixels)
#plot(samples)
#view(samples)
```

7.6.3 Performing NKDE

```{r}
densities <- nkde(network,
                  events = childcare,
                  w = rep(1, nrow(childcare)),
                  samples = samples,
                  kernel_name = "quartic",
                  bw = 300,
                  div= "bw",
                  method = "simple",
                  digits = 1,
                  tol = 1,
                  grid_shape = c(1,1),
                  max_depth = 8,
                  agg = 5,
                  sparse = TRUE,
                  verbose = FALSE)
```

7.6.3.1 Visualising NKDE Code chunk below will be used to insert the computed density values (i.e. densities) into samples and lixels objects as density field.

```{r}
samples$density <- densities
lixels$density <- densities
```

```{r}
# rescaling to help the mapping metre to kilometre
samples$density <- samples$density*1000
lixels$density <- lixels$density*1000
```

```{r}
tmap_mode('view')
tm_shape(lixels)+
  tm_lines(col="density")+
tm_shape(childcare)+
  tm_dots()
tmap_mode('plot')
```

7.7 Network Constrained G- and K-Function Analysis

```{r}
kfun_childcare <- kfunctions(network,
                             childcare,
                             start = 0,
                             end = 1000,
                             step = 50,
                             width = 50,
                             nsim = 50,
                             resolution = 50,
                             verbose = FALSE,
                             conf_int = 0.05)
```

```{r}
kfun_childcare$plotk
```
