---
title: "In-class Exercise 1"
author: "Chen Jinghui"
date: "Aug 26 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  message: false #no warning messages go out
  freeze: true  #will not re-render all items
---

**Setup**

Load Packages

```{r}
pacman::p_load(sf, tidyverse, tmap, ggstatsplot)
```

Importing polygon feature data in shapefile format

```{r}
mpsz14_shp = st_read(dsn = "C:/jhui-chen/ISSS626-GAA/In-Class_Ex/In-Class_Ex01/data",
layer = "MP14_SUBZONE_WEB_PL")

class(mpsz14_shp)
```

Importing GIS data in kml format KML file is not working...

```{r}
#| eval: false
mpsz14_kml = st_read("C:/jhui-chen/ISSS626-GAA/In-Class_Ex/In-Class_Ex01/data/MasterPlan2014SubzoneBoundaryWebKML.kml")
```

Using existing shpfile to convert back to kml format

```{R}
st_write(mpsz14_shp,
        "C:/jhui-chen/ISSS626-GAA/In-Class_Ex/In-Class_Ex01/data/MasterPlan2014SubzoneBoundaryWebKML.kml", 
        delete_dsn = TRUE)
```

Importing polygon feature data in shapefile format

```{r}
mpsz19_shp = st_read(dsn = "C:/jhui-chen/ISSS626-GAA/In-Class_Ex/In-Class_Ex01/data",
layer = "MPSZ-2019")

class(mpsz19_shp)
```

Importing GIS data in kml format. Note that the tidyiness of this dataset is not the same as the shp file version, it is bad.

```{r}
#| eval: false
mpsz19_kml = st_read("C:/jhui-chen/ISSS626-GAA/In-Class_Ex/In-Class_Ex01/data/MasterPlan2019SubzoneBoundaryNoSeaKML.kml")
```

Importing polygon feature data in shapefile format. The CRS (coordinate reference system is in WGS84, in decimal degree systems) This is a geographic coordinate system. Must check before doing distance analysis

```{r}
mpsz19_shp = st_read(dsn = "C:/jhui-chen/ISSS626-GAA/In-Class_Ex/In-Class_Ex01/data",
layer = "MPSZ-2019")
```

Transform coordinate system The %\>% allows you to run the two functions without creating an intermediate dataset. Will now be in svy21 format

```{r}
mpsz19_shp <- st_read(dsn = "data/",
                      layer = "MPSZ-2019") %>%
  st_transform(crs = 3414)
st_geometry(mpsz19_shp)
```
